A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2015-11-08 imaging_demo/ms.tex
Avoid Type 3 fonts in all figures for MNRAS specs
From dca9220 (mnras):
====================

(*) Remake all figures to not use Type 3 fonts to be
MNRAS-compatible

In ms.tex,

(*) Use updated figures
(*) Change use of commented F-engine figure to parallel_F_engine

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

@@

Fig.~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging. The
propagated electric fields are shown on the left at different time stamps,
$t_1\ldots t_\textrm{M}$. At each time stamp, the electric fields measured by
antennas are denoted by $E_1(t)\ldots E_\textrm{N}(t)$. The F-engine performs a
temporal Fourier transform on the electric field time-series to obtain electric
field spectra $\widetilde{E}_1(f)\ldots \widetilde{E}_\textrm{N}(f)$ 
($\widetilde{\mathbf{E}}(\mathbf{a})$ in matrix notation) for each of the 
antennas. Each of the complex antenna gains are calibrated to correct the 
corresponding electric field spectra. These calibrated electric fields are 
gridded using an antenna-based gridding convolution function after which it is 
spatially Fourier transformed and squared to obtain an image cube for every time 
stamp. These images are then time-averaged to obtain the accumulated image 
$I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).
\begin{figure}
  \includegraphics[width=\columnwidth]{figure1}
  % \includegraphics[width=\columnwidth]{MOFF_flowchart}
  % \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
  \caption{A flowchart of MOFF imaging in EPIC. The propagated electric fields
    shown on the left are measured as time-series $E_1(t)\ldots E_\textrm{N}(t)$
    by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then 
    Fourier transformed by the F-engine to produce electric field spectra 
    $\widetilde{E}_1(f)\ldots \widetilde{E}_\textrm{N}(f)$. They are calibrated 
    and gridded. The gridded electric fields $\widetilde{E}_\textrm{g}(f)$ from 
    each time series are imaged to produce images 
    $I_1(f)\ldots I_\textrm{M}(f)$. These images are time-averaged to obtain the 
    final image $I(f)$.}
  \label{fig:MOFF-flowchart}
\end{figure}

Fig.~\ref{fig:FX-flowchart} shows the flowchart for a visibility-based software 
holographic imaging from a FX correlator. The antenna-based F-engine is identical 
to that in the MOFF processing. The electric field spectra from each antenna are 
then cross-multiplied in the X-engine with those from all other antennas to 
obtain the visibilities $\widetilde{V}_\textrm{p}(f)$ ($\mathbf{m}(\mathbf{v})$ 
in matrix notation). They are calibrated and time-averaged to obtain 
$\langle \widetilde{V}_\textrm{p}(f)\rangle$ which are then gridded and imaged to 
obtain the image $I(f)$. The $I(f)$ obtained from both techniques are 
theoretically identical as explained in \S\ref{sec:math}.
\begin{figure}
  \includegraphics[width=\columnwidth]{figure2}
  % \includegraphics[width=\columnwidth]{FX_flowchart}
  % \includegraphics[width=\columnwidth]{FX_flowchart.eps}
  \caption{A flowchart of visibility-based software holographic imaging in EPIC. 
    The FX process flow shares the F-engine with the MOFF process. Following the 
    F-engine, the electric fields pass through the X-engine to produce 
    visibilities $\widetilde{V}_\textrm{p}(f)$ which are calibrated and 
    time-averaged. Then they are gridded to obtain the gridded visibilities 
    $\widetilde{V}_\textrm{g}(f)$ which are then Fourier transformed to obtain 
    the image, $I(f)$.}
  \label{fig:FX-flowchart}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figure4}
  % \includegraphics[width=\columnwidth]{F-engine}
  % \includegraphics[width=\columnwidth]{F-engine.eps}
  \caption{Block diagram of a F-engine. The electric field data streams from
    antennas are Fourier transformed in parallel to generate electric field
    spectra.}
  \label{fig:f-engine}
\end{figure}

@@

The different antennas are color coded by roughly the class of antenna size 
they fall into. Thus symbols of one color falling into a wedge of the same color 
indicates MOFF imaging is favoured for those telescopes and vice versa. For 
e.g., MOFF imaging is favoured in HERA-331 and HERA-6769 because they are inside 
the wedge but not so in cases of HERA-19 and HERA-37. A majority of the 
next-generation radio telescopes, namely, HERA-331 and its future expanded 
versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF 
imaging will be desirable. LWA1 and LWA-OV are already very close to the dividing 
line. Their hypothetical expansions,\footnote{LWA1-x2x1 and LWA1-x4x1.5, and 
LWA-OVx2x1 and LWA-OVx4x1.5 denote two-fold and four-fold increase in the number 
of antennas within a core diameter that is 1 and 1.5 times the current size of 
100~m and 200~m respectively for LWA1 and LWA-OV.} will be in the computational 
regime favoured by MOFF. For a fixed baseline length, regions favouring the MOFF 
algorithm tend to be towards large $\Nant$ indicating large-N dense array 
layouts with smaller antenna elements are best suited for deploying EPIC.

>>>>

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
% \pdfminorversion=5

@@

Fig.~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging. The
propagated electric fields are shown on the left at different time stamps,
$t_1\ldots t_\textrm{M}$. At each time stamp, the electric fields measured by
antennas are denoted by $E_1(t)\ldots E_\textrm{N}(t)$. The F-engine performs a
temporal Fourier transform on the electric field time-series to obtain electric
field spectra $\widetilde{E}_1(f)\ldots \widetilde{E}_\textrm{N}(f)$ 
($\widetilde{\mathbf{E}}(\mathbf{a})$ in matrix notation) for each of the 
antennas. Each of the complex antenna gains are calibrated to correct the 
corresponding electric field spectra. These calibrated electric fields are 
gridded using an antenna-based gridding convolution function after which it is 
spatially Fourier transformed and squared to obtain an image cube for every time 
stamp. These images are then time-averaged to obtain the accumulated image 
$I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).

\begin{figure}
  \includegraphics[width=\columnwidth]{figure1}
  % \includegraphics[width=\columnwidth]{MOFF_flowchart}
  % \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
  \caption{A flowchart of MOFF imaging in EPIC. The propagated electric fields
    shown on the left are measured as time-series $E_1(t)\ldots E_\textrm{N}(t)$
    by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then 
    Fourier transformed by the F-engine to produce electric field spectra 
    $\widetilde{E}_1(f)\ldots \widetilde{E}_\textrm{N}(f)$. They are calibrated 
    and gridded. The gridded electric fields $\widetilde{E}_\textrm{g}(f)$ from 
    each time series are imaged to produce images 
    $I_1(f)\ldots I_\textrm{M}(f)$. These images are time-averaged to obtain the 
    final image $I(f)$.}
  \label{fig:MOFF-flowchart}
\end{figure}

Fig.~\ref{fig:FX-flowchart} shows the flowchart for a visibility-based software 
holographic imaging from a FX correlator. The antenna-based F-engine is identical 
to that in the MOFF processing. The electric field spectra from each antenna are 
then cross-multiplied in the X-engine with those from all other antennas to 
obtain the visibilities $\widetilde{V}_\textrm{p}(f)$ ($\mathbf{m}(\mathbf{v})$ 
in matrix notation). They are calibrated and time-averaged to obtain 
$\langle \widetilde{V}_\textrm{p}(f)\rangle$ which are then gridded and imaged to 
obtain the image $I(f)$. The $I(f)$ obtained from both techniques are 
theoretically identical as explained in \S\ref{sec:math}.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure2}
  % \includegraphics[width=\columnwidth]{FX_flowchart}
  % \includegraphics[width=\columnwidth]{FX_flowchart.eps}
  \caption{A flowchart of visibility-based software holographic imaging in EPIC. 
    The FX process flow shares the F-engine with the MOFF process. Following the 
    F-engine, the electric fields pass through the X-engine to produce 
    visibilities $\widetilde{V}_\textrm{p}(f)$ which are calibrated and 
    time-averaged. Then they are gridded to obtain the gridded visibilities 
    $\widetilde{V}_\textrm{g}(f)$ which are then Fourier transformed to obtain 
    the image, $I(f)$.}
  \label{fig:FX-flowchart}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figure4}
  % \includegraphics[width=\columnwidth]{parallel_F_engine}
  % \includegraphics[width=\columnwidth]{parallel_F_engine.eps}
  \caption{Block diagram of a F-engine. The electric field data streams from
    antennas are Fourier transformed in parallel to generate electric field
    spectra.}
  \label{fig:f-engine}
\end{figure}

@@

The different antennas are color coded by roughly the class of antenna size 
they fall into. Thus symbols of one color falling into a wedge of the same color 
indicates MOFF imaging is favoured for those telescopes and vice versa. For 
e.g., MOFF imaging is favoured in HERA-331 and HERA-6769 because they are inside 
the wedge but not so in cases of HERA-19 and HERA-37. A majority of the 
next-generation radio telescopes, namely, HERA-331 and its future expanded 
versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF 
imaging will be desirable. LWA1 and LWA-OV are already very close to the dividing 
line. Their hypothetical expansions,\footnote{LWA1-x2x1 and LWA1-x4x1.5, and 
LWA-OVx2x1 and LWA-OVx4x1.5 denote two-fold and four-fold increase in number 
of antennas within a core diameter that is 1 and 1.5 times the current size of 
100~m and 200~m respectively for LWA1 and LWA-OV.} will be in the computational 
regime favoured by MOFF. For a fixed baseline length, regions favouring the MOFF 
algorithm tend to be towards large $\Nant$ indicating large-N dense array 
layouts with smaller antenna elements are best suited for deploying EPIC.
